Staged Abstract Interpreters

### 2019-03-28 paper/20_prelim.tex
misc and reorganize
fix a mistake

We also defines several methods to obtain and modify the environment and store.
These methods either returns monadic values with environment or store inside, or
simply a @Unit@ value for effects. For example, @ext_env@ takes a value @ans@ of
type @Ans@ and a pair of binding that maps @x@ to $\alpha$, and installs this
binding to the environment when evaluating @ans@.

@@

  type AnsM[T] <: MonadOps[R, AnsM, T]
  type Ans = AnsM[Value]
    \end{lstlisting}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.55\textwidth}
    \begin{lstlisting}
// Environment operations
def ask_env: AnsM[Env]
def ext_env(ans: Ans)(xα: (Ident, R[Addr])): Ans
// Store operations
def get_store: AnsM[Store]
def put_store(σ: R[Store]): AnsM[Unit]
def set_store(av: (R[Addr], R[Value])): AnsM[Unit]
    \end{lstlisting}
  \end{subfigure}
\end{figure}
\vspace{-1em}

@@

\vspace{-1em}
\begin{figure}[h!]
  \centering
  \begin{lstlisting}
            def eval(ev: Expr => Ans)(e: Expr): Ans = e match {
              case Lit(i) => num(i)                    case Let(x, rhs, e) => for {
              case Var(x) => for {                       v <- ev(rhs)
                ρ <- ask_env                             α <- alloc(x)
                σ <- get_store                           _ <- set_store(α → v)
              } yield get(σ, ρ, x)                      rt <- ext_env(ev(e))(x → α)
              case Lam(x, e) => for {                  } yield rt
                ρ <- ask_env                           case Aop(op, e1, e2) => for {
              } yield close(ev)(Lam(x, e), ρ)            v1 <- ev(e1)                                               
              case App(e1, e2) => for {                  v2 <- ev(e2)
                v1 <- ev(e1)                           } yield arith(op, v1, v2)
                v2 <- ev(e2)                           case Rec(x, rhs, e) => for {
                rt <- ap_clo(ev)(v1, v2)                 α <- alloc(x)
              } yield rt                                 v <- ext_env(ev(rhs))(x → α)
              case If0(e1, e2, e3) => for {              _ <- set_store(α → v)
                cnd <- ev(e1)                            rt <- ext_env(ev(e))(x → α)
                rt <- br0(cnd, ev(e2), ev(e3))         } yield rt                    
              } yield rt
            }
  \end{lstlisting}
\caption{The generic interpreter,
  shared by the unstaged/staged + concrete/abstract interpreter.}
\label{fig:shared_int}
\end{figure}
\vspace{-1em}

@@

\subsection{A Concrete Interpreter} \label{unstaged_conc}

@@

\begin{lstlisting}
  def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = rator match {
    case CloV(Lam(x, e), ρ: Env) => for {
      α <- alloc(x)
      _ <- set_store(α → rand)
      rt <- ext_env(ev(e))(x → α)
    } yield rt
  }
\end{lstlisting}

>>>>

We also defines several methods to obtain and modify the environment and store.
These methods either returns monadic values with environment or store inside, or
simply a @Unit@ value for effects. For example, @local_env@ takes a value @ans@ of
type @Ans@ and a new environment which will be installed when evaluating @ans@.

@@

  type AnsM[T] <: MonadOps[R, AnsM, T]
  type Ans = AnsM[Value]
    \end{lstlisting}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.55\textwidth}
    \begin{lstlisting}
// Environment operations
def ask_env: AnsM[Env]
def local_env(ans: Ans)(ρ: R[Env]): Ans
// Store operations
def get_store: AnsM[Store]
def put_store(σ: R[Store]): AnsM[Unit]
def set_store(av: (R[Addr], R[Value])): AnsM[Unit]
    \end{lstlisting}
  \end{subfigure}
\end{figure}
\vspace{-1em}

@@

\vspace{-1em}
\begin{figure}[h!]
  \centering
  \begin{lstlisting}
            def eval(ev: Expr => Ans)(e: Expr): Ans = e match {
              case Lit(i) => num(i)                   case Let(x, rhs, e) => for {
              case Var(x) => for {                       v <- ev(rhs)
                ρ <- ask_env                             ρ <- ask_env
                σ <- get_store                           α <- alloc(x)
              } yield get(σ, ρ, x)                       _ <- set_store(α → v)
              case Lam(x, e) => for {                   rt <- local_env(ev(e))(ρ + (x → α))
                ρ <- ask_env                          } yield rt
              } yield close(ev)(Lam(x, e), ρ)         case Aop(op, e1, e2) => for {
              case App(e1, e2) => for {                 v1 <- ev(e1)                                               
                v1 <- ev(e1)                            v2 <- ev(e2)
                v2 <- ev(e2)                          } yield arith(op, v1, v2)
                rt <- ap_clo(ev)(v1, v2)              case Rec(x, rhs, e) => for {
              } yield rt                                α <- alloc(x)
              case If0(e1, e2, e3) => for {             ρ <- ask_env
                cnd <- ev(e1)                           v <- local_env(ev(rhs))(ρ + (x → α))
                rt <- br0(cnd, ev(e2), ev(e3))          _ <- set_store(α → v)
              } yield rt                                rt <- local_env(ev(e))(ρ + (x → α))
                                                      } yield rt                    
            }
  \end{lstlisting}
\caption{The generic interpreter,
  shared by the unstaged/staged + concrete/abstract interpreter.}
\label{fig:shared_int}
\end{figure}
\vspace{-1em}

@@

\section{A Concrete Interpreter} \label{unstaged_conc}

@@

\begin{lstlisting}
  def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = rator match {
    case CloV(Lam(x, e), ρ: Env) => for {
      α <- alloc(x)
      _ <- set_store(α → rand)
      rt <- local_env(ev(e))(ρ + (x → α))
    } yield rt
  }
\end{lstlisting}

### 2019-03-28 paper/40_absinterp.tex
fix a mistake

\begin{lstlisting}
  def lift_nd[T](vs: List[T]): AnsM[T]
  def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = for {
    α <- alloc(x)
    CloV(Lam(x, e), ρ: Env) <- lift_nd[AbsValue](rator.toList)
    _ <- set_store(α → rator)
    rt <- ext_env(ev(e))(x → α)
  } yield rt
\end{lstlisting}

>>>>

\begin{lstlisting}
  def lift_nd[T](vs: List[T]): AnsM[T]
  def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = for {
    CloV(Lam(x, e), ρ) <- lift_nd[AbsValue](rator.toList)
    α <- alloc(x)
    _ <- set_store(α → rator)
    rt <- local_env(ev(e))(ρ + (x → α))
  } yield rt
\end{lstlisting}
